Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-11 ms.tex
random stuff3

\begin{figure}
  \includegraphics[width=\columnwidth]{profiles.eps}
  \caption{\label{fig:profiles}Radial profiles of the CNM density
    ({\it top}), temperature ({\it middle}), and velocity ({\it bottom}),
    calculated for a representative sample of galaxies.  Different colors
    denote different values of the effective wind heating rate,
    $\vwO=1200$ km s$^{-1}$ ({\it blue}), $\vwO$=600 km s$^{-1}$ ({\it
      orange}), and 300 km s$^{-1}$ ({\it green}).  Different
    linestyles correspond to different values of $\Mbh$, $\Mbh=10^6
    \Msun$ ({\it dot-dashed}), $\Mbh=10^7 \Msun$ ({\it solid}), and
    $\Mbh=10^8 \Msun$ ({\it dashed}). Thin lines correspond to cusp
    galaxies ($\Gamma$=0.8) and thick lines correspond to core
    galaxies ($\Gamma$=0.1). Squares mark the location of
    the stagnation radius for each solution
 }
\end{figure}

@@

The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 27 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 30 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  We caveat that, unlike
stellar wind heating, Compton heating depends on radius, scaling as
$v_{w}^{\rm C}(r) \propto (n^{2}\xi/\rho_{\star})^{1/2} \propto
r^{(\Gamma-\densSlope-1)/2}$, i.e. $\propto r^{-0.6}$ and $\propto r^{-0.75}$
for core ($\Gamma = 0.8$; $\densSlope \simeq 1$) and cusp ($\Gamma = 0.1$;
$\densSlope \simeq 0.6$) galaxies, respectively.  Our model's assumption that the heating
parameter be radially constant is not satisfied, but this variation is sufficiently weak that it should not significantly alter
our conclusions.

If Compton heating acts alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then solving equation
(\ref{eq:vC}) for $\tilde{v}_{\rm w}$ yields
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 150 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 96 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
Compton heating can thus be significant in young stellar populations with relatively high mass loss rates, e.g. $v_{w}^{\rm C} \gtrsim 300$ km s$^{-1}$ for $\eta \gtrsim 1$.  The mass accretion rate corresponding to the state where Compton heating limits the average accretion rate is given by substituting equation (\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 3\times 10^{-3} \eta_{0.02}^{0.20}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.15}
&, \text{core}\\ 8\times 10^{-4} \eta_{0.02}^{0.30}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp},
  \end{cases}
  \label{eq:MdotCa}
\end{align}
If the accretion luminosity $\epsilon = 10 \dot{M}/\dot{M}_{\rm edd}$
depends on the Eddington ratio as predicted by some models for
radiatively inefficient accretion flows (\citealt{Narayan&Yi95};
\citealt{Narayan+98}; \citealt{XieYuan:2012a}), then the Compton-limited accretion rate becomes
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 1.8\times 10^{-2} \eta_{0.02}^{0.11}T_{\rm
C,9}^{-0.44}M_{\bullet,8}^{0.08}
&, \text{core}\\ 8.7\times 10^{-4} \eta_{0.02}^{0.17}T_{\rm
C,9}^{-0.4} M_{\bullet,8}^{0.08}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotC}
\end{align}
{\bf  Total efficiency is 10 (mdot/MdotEdd). Everywhere.}

@@

Assume that the outflow power is proportional to the SMBH accretion rate, $L_{\rm j} =
\epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is a jet efficiency factor.  Further assume that this energy is deposited as heat uniformly interior to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm
heat}^{3}$.  The resulting volumetric heating rate $\dot{e} = \epsilon_{\rm
j}\dot{M}c^{2}/V_{\rm heat}$ near the stagnation radius results in a heating parameter given by
\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}|_{r_{s}}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray} 
where we have used the facts that $\dot{M} = \eta M_{\star}|_{r_{\rm s}}/t_{\rm h}$ and, for radii inside the Nuker break radius, $M_{\star}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ (eq.~[\ref{eq:rhostar}]).  If the bulk of the outflow energy is released near the stagnation radius, then even a small heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other sources of non-accretion powered heating.  However, if this energy is deposited over much larger physical scales comparable to the size of the galaxy, i.e. $r_{\rm heat} \gtrsim $ 10 kpc $\sim 10^{4}r_{\rm s}$, then kinetic feedback is unimportant, even for a powerful jet with $\epsilon_{\rm j} \sim 1$.

@@

As in the case of Compton heating, kinetic heating could in principle  `self-regulate' the accretion flow insofar as a lower heating rate results in a higher accretion rate (eq.~[\ref{eq:vjet}]), which in turn may create stronger jet feedback.  However, as it is not possible to make concrete predictions for such a process given the uncertainty in the efficiency of jet heating (see $\S\ref{sec:heating}$ for further discussion).

@@

Figure~\ref{fig:vwSourcesImp} shows the contributions of various heating sources as a function of time $\tau_{\star}$ after a burst of star formation for black holes of mass $\Mbh = 10^{6} \Msun$ (top panel) and $\Mbh = 10^{8} \Msun$ (bottom panel). The heating and mass input parameters due to stellar winds, $v_{w}^{\star}(\tau_{\star})$ and $\eta(\tau_{\star})$, are calculated as described in Appendix~\ref{app:windheat} (Fig.~\ref{fig:vwImp}).  The heating from SN Ia ($v_{w}^{\rm Ia}$) and Compton heating ($v_{\rm w}^{\rm C}$) are calculated from equations~\eqref{eq:vIa} and~\eqref{eq:vC} respectively, also making use of $\eta(\tau_{\star})$ from stellar winds.  SN Ia heating depends also on a convolution of the DTD (eq. [\ref{eq:DTD}]) and the star formation history, which reduces to the DTD itself for a single star burst.  We crudely account for the expected suppression of SN Ia heating resulting from its non-steady nature by attenuating $v_{w}^{\rm Ia}$ by a factor $e^{-\rs/\rIa}$, where $r_{\rm Ia}$ is the radius exterior to which the time between subsequent supernovae exceeds the local dynamical time (eq.~[\ref{eq:rIa}]).  Finally, note that $v_{\rm w}^{\rm C}$ and $v_{\rm w}^{\rm Ia}$ also depend on the total heating rate (eq.~[\ref{eq:vtot}]), requiring us to simultaneously solve a series of implicit equations to determine each heating rate.  Heating from MSPs is found to be negligibly small for fiducial parameters and hence is not shown.

>>>>

\begin{figure}
  \includegraphics[width=\columnwidth]{profiles.eps}
  \caption{\label{fig:profiles}Radial profiles of the CNM density
    ({\it top}), temperature ({\it middle}), and velocity ({\it bottom}),
    calculated for a representative sample of galaxies.  Different colors
    denote different values of the effective wind heating rate,
    $\vwO=1200$ km s$^{-1}$ ({\it blue}), $\vwO$=600 km s$^{-1}$ ({\it
      orange}), and 300 km s$^{-1}$ ({\it green}).  Different
    linestyles correspond to different values of $\Mbh$, $\Mbh=10^6
    \Msun$ ({\it dot-dashed}), $\Mbh=10^7 \Msun$ ({\it solid}), and
    $\Mbh=10^8 \Msun$ ({\it dashed}). Thin lines correspond to cusp
    galaxies ($\Gamma$=0.8) and thick lines correspond to core
    galaxies ($\Gamma$=0.1). Squares mark the location of
    the stagnation radius for each solution.
 }
\end{figure}

@@

The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon \dot{M}c^{2}$, where
$\epsilon$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 27 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 30 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon/0.01 \sim 1$.  We caveat that, unlike
stellar wind heating, Compton heating depends on radius, scaling as
$v_{w}^{\rm C}(r) \propto (n^{2}\xi/\rho_{\star})^{1/2} \propto
r^{(\Gamma-\densSlope-1)/2}$, i.e. $\propto r^{-0.6}$ and $\propto r^{-0.75}$
for core ($\Gamma = 0.8$; $\densSlope \simeq 1$) and cusp ($\Gamma = 0.1$;
$\densSlope \simeq 0.6$) galaxies, respectively.  Our model's assumption that the heating
parameter be radially constant is not satisfied, but this variation is sufficiently weak that it should not significantly alter
our conclusions.

If Compton heating acts alone, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then solving equation
(\ref{eq:vC}) for $\tilde{v}_{\rm w}$ yields
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 150 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 96 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
Compton heating can thus be significant in young stellar populations with relatively high mass loss rates, e.g. $v_{w}^{\rm C} \gtrsim 300$ km s$^{-1}$ for $\eta \gtrsim 1$.  The mass accretion rate corresponding to the state where Compton heating limits the average accretion rate is given by substituting equation (\ref{eq:vC2}) into equation (\ref{eq:eddr_analytic}):
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 3\times 10^{-3} \eta_{0.02}^{0.20}T_{\rm
C,9}^{-0.8}\epsilon_{-2}^{-0.8} M_{\bullet,8}^{0.15}
&, \text{core}\\ 8\times 10^{-4} \eta_{0.02}^{0.30}T_{\rm
C,9}^{-0.7}\epsilon_{-2}^{-0.7} M_{\bullet,8}^{0.14}
&, \text{cusp},
  \end{cases}
  \label{eq:MdotCa}
\end{align}
If the accretion luminosity $\epsilon = 10 \dot{M}/\dot{M}_{\rm edd}$
depends on the Eddington ratio as predicted by some models for
radiatively inefficient accretion flows (\citealt{Narayan&Yi95};
\citealt{Narayan+98}; \citealt{XieYuan:2012a}), then the Compton-limited accretion rate becomes
\begin{align}
\frac{\dot{M}_{C}}{\dot{M}_{\rm edd}} \approx 
\begin{cases} 1.8\times 10^{-2} \eta_{0.02}^{0.11}T_{\rm
C,9}^{-0.44}M_{\bullet,8}^{0.08}
&, \text{core}\\ 8.7\times 10^{-4} \eta_{0.02}^{0.17}T_{\rm
C,9}^{-0.4} M_{\bullet,8}^{0.08}
&, \text{cusp}.
  \end{cases}
  \label{eq:MdotC}
\end{align}
%{\bf  Total efficiency is 10 (mdot/MdotEdd). Everywhere.}

@@

Assume that the outflow power is proportional to the SMBH accretion rate, $L_{\rm j} =
\epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is a jet efficiency factor.  Further assume that this energy is deposited as heat uniformly interior to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm
heat}^{3}$.  The resulting volumetric heating rate $e = \epsilon_{\rm
j}\dot{M}c^{2}/V_{\rm heat}$ near the stagnation radius results in a heating parameter given by
\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}|_{r_{s}}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray} 
where we have used the facts that $\dot{M} = \eta M_{\star}|_{r_{\rm s}}/t_{\rm h}$ and, for radii inside the Nuker break radius, $M_{\star}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ (eq.~[\ref{eq:rhostar}]).  If the bulk of the outflow energy is released near the stagnation radius, then even a small heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other sources of non-accretion powered heating.  However, if this energy is instead deposited over much larger physical scales comparable to the size of the galaxy, i.e. $r_{\rm heat} \gtrsim $ 10 kpc $\sim 10^{4}r_{\rm s}$, then kinetic feedback is unimportant, even for a powerful jet with $\epsilon_{\rm j} \sim 1$.

@@

As in the case of Compton heating, kinetic heating could in principle  `self-regulate' the accretion flow insofar as a lower heating rate results in a higher accretion rate (eq.~[\ref{eq:vjet}]), which in turn may create stronger jet feedback.  However, given the uncertainty in the efficiency of jet heating, we defer a more detailed discussion to $\S\ref{sec:kinetic}$.

@@

Figure~\ref{fig:vwSourcesImp} shows the contributions of various heating sources as a function of time $\tau_{\star}$ after a burst of star formation for black holes of mass $\Mbh = 10^{6} \Msun$ (top panel) and $\Mbh = 10^{8} \Msun$ (bottom panel). The heating and mass input parameters due to stellar winds, $v_{w}^{\star}(\tau_{\star})$ and $\eta(\tau_{\star})$, are calculated as described in Appendix~\ref{app:windheat} (Fig.~\ref{fig:vwImp}).  The heating from SN Ia ($v_{w}^{\rm Ia}$) and Compton heating ($v_{\rm w}^{\rm C}$) are calculated from equations~\eqref{eq:vIa} and~\eqref{eq:vC} respectively, also making use of $\eta(\tau_{\star})$ from stellar winds.  SN Ia heating depends also on a convolution of the DTD (eq. [\ref{eq:DTD}]) and the star formation history, which reduces to the DTD itself for a single star burst.  We crudely account for the expected suppression of SN Ia heating resulting from its non-steady nature by attenuating $v_{w}^{\rm Ia}$ by a factor $e^{-\rs/\rIa}$, where $r_{\rm Ia}$ is the radius exterior to which the time between subsequent supernovae exceeds the local dynamical time (eq.~[\ref{eq:rIa}]).  Finally, note that $v_{\rm w}^{\rm C}$ and $v_{\rm w}^{\rm Ia}$ also depend on the total heating rate (eq.~[\ref{eq:vtot}]), requiring us to simultaneously solve a series of implicit equations to determine each heating rate.  Heating from MSPs is found to be negligibly small for fiducial parameters and hence is not shown, while kinetic feedback is neglected given its uncertain efficiency ($\S\ref{sec:kinetic}$).
